Results

### 2021-08-11 Manuscript/Manuscript.Rmd
ko

@@

Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018]. Binary logistic regression is used for predicting binary classes, such gender of Palmer amaranth plants. The whole dataset was splitted into 80% train and 20% test data. The 80% train is used for the model training and the rest 20% is used for checking how the model generalized on unseen dataset. With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables. The model family was binomial with a logit function. The model fit was assessed through pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package). The marginal effects computation was performed with Average Marginal Effects at every observed value of x and average across the results (AMEs) [@leeper2017] using *margins* function from "margins" package. The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.

@@

## Cumulative flowering {-}

@@

## Height and biomass {-}

Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A). At first cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively. Moreover, Palmer amaranth produce less biomass at cohort time. Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.

The Palmer amaranth achieve 92.3 cm tall when growing at first cohort timing in corn (Figure 3B).

@@

Palmer amaranth grow and development at second cohort was limited due to the crop competitive ability at advanced development stages. Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness. As a result, Palmer amaranth height and biomass was lower compared to its first cohort. Under crop canopy, Palmer amaranth flowering pattern was near to similar in corn and soybean. Palmer amaranth growing without crop competition produced highest amounts of biomass and less extended flowering pattern. The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height. Nonetheless, Palmer amaranth produced 31% less biomass in second cohort compared to first cohort timing. In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015]. These results suggests that crop competition is not the only factor limiting late Palmer amaranth establishment. The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to thermal units (e.g, growing degree days). It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period. A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021]. In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.

Seed production was not evaluated due to Palmer amaranth harvest at flowering time. Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018]. I our study, Palmer amaranth growing at first cohort accumulated an overall 50% more biomass when compared to second cohort. Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem. Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015]. Nonetheless, seed production at second cohort is likely to replenish the soil seedbank. Seed production and deposition in the seedbank is a key factor for species perpetuation [@menges1987]. Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].

Palmer amaranth flowering pattern was variable

Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems. Palmer amaranth management priority should focus on minimizing Palmer amaranth dispersal into new geographies. Early-season management programs would have a large negative effect in Palmer amaranth growth and development. Tactics that promote crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g. cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.

>>>>

Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018]. Binary logistic regression is used for predicting binary classes, such gender of Palmer amaranth plants. Prior to the analysis, all missing values were removed from the dataset. Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location. The whole dataset was splitted into 80% train and 20% test data. The 80% train is used for the model training and the rest 20% is used for checking how the model generalized on unseen dataset. With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables. The model family was binomial with a logit function. The model fit was assessed through pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package). The marginal effects computation was performed with Average Marginal Effects at every observed value of x and average across the results (AMEs) [@leeper2017] using *margins* function from "margins" package. The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function. The model quality prediction from the classification algorithm was measure with precision (*precision* function), recall (*recall* function) and F1 score (*f_meas* function) using "yardstick" package. The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1 score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0. F1 score conveys the balance between the precision and the recall.

@@

@@

## Palmer amaranth cumulative flowering {-}

@@

## Palmer amaranth gender {-}

The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively. The likelihood ratio test showed a p-value of < 0.00. The marginal effects showed that increasing a weight unit by 1 g increases the probability of having a female plant by 0.08%. Similar trend is observed to height as well as doy, whereas the probability of being female increase by 0.16% and 0.4% when a unit of height (cm) and day increases.

The model evaluation accuracy in the unseen 20% dataset was 0.61 with a cutoff value for female and male plants of 0.49. The model classification showed a precision of 0.59, recall of 0.76, and a F1 score of 0.66. In addition, the area under the curve was 0.61.

## Palmer amaranth height and biomass {-}

Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A). At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.

Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74). In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75). At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04). The tallest and smallest Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2 cm tall at first cohort and 38.2 cm at second cohort timings.

@@

Palmer amaranth grow and development at second cohort was limited due to the crop competitive ability at advanced development stages. Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness. As a result, Palmer amaranth height and biomass was lower compared to its first cohort. Under crop canopy, Palmer amaranth flowering pattern was near to similar in corn and soybean. Palmer amaranth growing without crop competition produced highest amounts of biomass and less extended flowering pattern. The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height. Nonetheless, Palmer amaranth produced 21% less biomass in second cohort compared to first cohort timing. In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015]. These results suggests that crop competition is not the only factor limiting late Palmer amaranth establishment. The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to thermal units (e.g, growing degree days). It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period. A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021]. In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.

Seed production was not evaluated due to Palmer amaranth harvest at flowering time. Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018]. I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort. Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem. Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015]. Nonetheless, seed production at second cohort is likely to replenish the soil seedbank. Seed production and deposition in the seedbank is a key factor for species perpetuation [@menges1987]. Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].

Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings. Palmer amaranth growing in fallow and corn resulted in the overall shortest and longest flowering pattern. When growing in soybean, Palmer amaranth flowering pattern was similar to bareground and corn agroecossystems at first and second cohort timings, respectively. It seems that Palmer amaranth cohorts tend to have an extended flowering pattern when in high competition. The agroecossystems influenced Palmer amaranth gender. When growing in corn, the probability of having Palmer amaranth female plants reduced 18.1%.

Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems. Palmer amaranth management priority should focus on minimizing Palmer amaranth dispersal into new geographies. Early-season management programs would have a large negative effect in Palmer amaranth growth and development. Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g. cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.

### 2021-08-11 Manuscript/Manuscript.tex
ko

@@

\hypertarget{cumulative-flowering}{%
\subsection*{Cumulative flowering}\label{cumulative-flowering}}
\addcontentsline{toc}{subsection}{Cumulative flowering}

@@

\hypertarget{height-and-biomass}{%
\subsection*{Height and biomass}\label{height-and-biomass}}
\addcontentsline{toc}{subsection}{Height and biomass}

Palmer amaranth accumulated more biomass when growing in fallow compared
to Palmer amaranth growing in soybean and corn (Figure 3A). At first
cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1
and 14.9 g plant-1 in fallow, soybean and corn, respectively. Moreover,
Palmer amaranth produce less biomass at cohort time. Palmer amaranth
growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean
and 1.4 g plant when growing in corn.

The Palmer amaranth achieve 92.3 cm tall when growing at first cohort
timing in corn (Figure 3B).

@@

@@

>>>>

Palmer amaranth gender was fitted to a binary logistic regression
(Bangdiwala, 2018). Binary logistic regression is used for predicting
binary classes, such gender of Palmer amaranth plants. Prior to the
analysis, all missing values were removed from the dataset. Also, data
from Grant was not used in this analysis due to the uniform plant
harvesting at that location. The whole dataset was splitted into 80\%
train and 20\% test data. The 80\% train is used for the model training
and the rest 20\% is used for checking how the model generalized on
unseen dataset. With 80\% dataset, a binary response variable, male (0)
and female (1), was fitted to a generalized linear model (\emph{glm}
function) including day of year harvest, height, weight, crop and month
as independent variables. The model family was binomial with a logit
function. The model fit was assessed through pseudo R squared values
(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
\emph{nagelkerke} function (``rcompanion'' package). The marginal
effects computation was performed with Average Marginal Effects at every
observed value of x and average across the results (AMEs) (Leeper, 2017)
using \emph{margins} function from ``margins'' package. The rest 20\%
dataset was predicted using \emph{predict} function with a cutoff
estimation for male or female using \emph{performance} function. The
model quality prediction from the classification algorithm was measure
with precision (\emph{precision} function), recall (\emph{recall}
function) and F1 score (\emph{f\_meas} function) using ``yardstick''
package. The precision determines the accuracy of positive predictions
(female plants), recall determines the fraction of positives that were
correctly identified, and F1 score is a weighted harmonic mean of
precision and recall with the best score of 1 and the worst score of 0.
F1 score conveys the balance between the precision and the recall.

@@

\hypertarget{palmer-amaranth-cumulative-flowering}{%
\subsection*{Palmer amaranth cumulative
flowering}\label{palmer-amaranth-cumulative-flowering}}
\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}

@@

\hypertarget{palmer-amaranth-gender}{%
\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
\addcontentsline{toc}{subsection}{Palmer amaranth gender}

The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from
McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
likelihood ratio test showed a p-value of \textless{} 0.00. The marginal
effects showed that increasing a weight unit by 1 g increases the
probability of having a female plant by 0.08\%. Similar trend is
observed to height as well as doy, whereas the probability of being
female increase by 0.16\% and 0.4\% when a unit of height (cm) and day
increases.

The model evaluation accuracy in the unseen 20\% dataset was 0.61 with a
cutoff value for female and male plants of 0.49. The model
classification showed a precision of 0.59, recall of 0.76, and a F1
score of 0.66. In addition, the area under the curve was 0.61.

\hypertarget{palmer-amaranth-height-and-biomass}{%
\subsection*{Palmer amaranth height and
biomass}\label{palmer-amaranth-height-and-biomass}}
\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}

Palmer amaranth accumulated more biomass when growing in fallow compared
to Palmer amaranth growing in soybean and corn (Figure 3A). At first
cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1
and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second
cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed
by 6.3 g plant in soybean, and 1.4 g plant in corn.

Palmer amaranth height was more uniform across cohort timings, except
when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2
cm tall when growing at first cohort timing in bareground, which was not
different to 70.7 cm tall at second cohort timing (P = 0.74). In
addition, no difference in Palmer amaranth height (69.3 cm) was found in
soybean at first cohort compared to fallow (P \textgreater{} 0.75). At
second cohort in soybean, Palmer amaranth was near to 10 cm lower
compared to its first cohort time (P = 0.04). The tallest and smallest
Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2
cm tall at first cohort and 38.2 cm at second cohort timings.

@@

\leavevmode\hypertarget{ref-bangdiwala2018}{}%
Bangdiwala, S. I. (2018). Regression: Binary logistic.
\emph{International Journal of Injury Control and Safety Promotion} 25,
336--338.
doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.

@@

\leavevmode\hypertarget{ref-leeper2017}{}%
Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
Marginal Eﬀects} with {R}'s margins. 31.

### 2021-08-11 data analysis/Analysis.Rmd
ko

>>>>

```{r}
library(popbio)
?logi.hist.plot()
```

```{r}
data2 %>% 
  mutate(gender_A = case_when(
    gender == "m" ~ TRUE,
    gender == "f" ~ FALSE,
    TRUE ~ NA
  )) %>% 
  mutate_if(is_character, as.factor) %>% 
  filter(!is.na(gender_A)) %>% 
  mutate(doyh = as.numeric(doyh)) -> data3
```

```{r}
logi.hist.plot(data3$weight, data3$gender_A, 
               logi.mod = 1, boxp=FALSE, counts=FALSE, int= 10,
               type = "hist", rug = FALSE, col="gray")

ggsave("fig.png")
```

```{r}
aq.trans$survived <- aq.trans$fate!="dead"
a <- subset(aq.trans, leaf<50 & stage!="recruit", c(leaf,survived))
logi.hist.plot(a$leaf,  a$survived,
  type="hist", boxp=FALSE, counts=TRUE, int=10,
  ylabel="Survival probability", ylabel2="Number of plants",
  xlab="Number of leaves")
b <- glm(survived ~ leaf, binomial, data=a)
summary(b)
```

```{r}
data2 %>% 
  mutate(gender_A = case_when(
    gender == "m" ~ 0,
    gender == "f" ~ 1,
    TRUE ~ NA_real_
  )) %>% 
  mutate_if(is_character, as.factor) %>% 
  filter(!is.na(gender_A)) %>% 
  filter(!is.na(doyh)) %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(height)) %>% 
  mutate(doyh = as.numeric(doyh)) -> data3
```

```{r}
seq(min(mtcars$hp), max(mtcars$hp),len=100)
```

```{r}
fit = glm(gender_A ~ doyh, data = data3, family=binomial)
newdat <- data.frame(doyh=seq(min(data3$doyh), max(data3$doyh), len=length(data3$doyh)))
newdat$vs = predict(fit, newdata=newdat, type="response")
plot(gender_A ~ doyh, data=data3, col="red4")
lines(gender_A ~ doyh, newdat, col="green4", lwd=2)
```

```{r}
label1 <- tibble(ed = "10% flowering", 
                   x = 2.35, y = 156.5, label = "June 01")
rect1 <- tibble(ed = "50% flowering", 
               x = 2.85, y = 195, xend = 2.7, yend = 199)

label2 <- tibble(ed = "10% flowering", 
                   x = 2.5, y = 186.5, label = "July 01")
rect2 <- tibble(ed = "50% flowering",  
               x = 3.05, y = 229, xend = 2.9, yend = 235)

month <- tibble(ed = "50% flowering",
                x = c(0.5, 0.5, 0.5, 0.5),
                y = c(168, 197, 228.5, 259), 
                label = c("June", "July", "August", "September"))

label3 <- tibble(ed = "50% flowering", 
                   x = 2.9, y = 240, label = "Second \ncohort")

label4 <- tibble(ed = "50% flowering", 
                   x = 2.7, y = 205, label = "First \ncohort")
```

```{r eval = FALSE, include=FALSE}
result3 %>% 
  mutate(ed = case_when(
    ed == "10" ~ "10% flowering",
    ed == "50" ~ "50% flowering",
    ed == "90" ~ "90% flowering",
    TRUE ~ "ed")) %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>% 
  ggplot(aes(x = crop, y = estimate)) +
  geom_segment(aes(x = 0.8, xend = 3.2, y = 153, yend = 153), 
             color = 1,
             alpha = 0.5) +
  geom_segment(aes(x = 0.8, xend = 3.2, y = 183, yend = 183), 
             color = 2,
             alpha = 0.5) +
  scale_x_discrete(expand = c(0.1, 0.1)) +
  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
                        ymin= 153, ymax= 183), 
            fill = "#FFFFF0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
                        ymin= 183, ymax= 214), 
            fill = "beige", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
                        ymin= 214, ymax= 244), 
            fill = "#F0FFF0", alpha=0.05) +
  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
                        ymin= 244, ymax= 274), 
            fill = "#F8F8FF", alpha=0.5) +
  geom_text(data = month, aes(x = x, y = y), label = month$label,
            size = 2, color = c("#8B8000", "#80461b", 
                                "darkgreen", "blue")) +
  scale_color_manual(values = c(1, 2)) +
  labs(x = NULL, y = "Day of year",
       title = NULL) +
  geom_curve(data = rect1, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             arrow = arrow(length = unit(0.07, "inch")), 
             color = 1,
             size = 0.4, curvature = 0.3) + 
  geom_curve(data = rect2, aes(x = x, y = y, 
                               xend = xend, yend = yend),
             color = 2,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = 0.3) + 
  geom_text(data = label1, 
            aes(x = x, y = y), 
            size = 2, hjust = 0, angle = 270,
            label = label1$label, color = 1) +
  geom_text(data = label2, 
            aes(x = x, y = y), 
            size = 2, hjust = 0, angle = 270,
            label = label2$label, color = 2) +
  geom_text(data = label3, 
            aes(x = x, y = y), 
            size = 2, hjust = 0, 
            label = label3$label, color = 2) +
  geom_text(data = label4, 
            aes(x = x, y = y), 
            size = 2, hjust = 0, 
            label = label4$label, color = 1) +
  facet_grid(~ ed, scales = "free_y") +
  geom_point(aes(color = month), 
             size = 1,
             position = position_dodge2(width = 0.4)) +
  geom_linerange(aes(ymin = lower, ymax = upper, color = month), 
                 size = 1.3, position = position_dodge2(width = 0.4)) +
  coord_flip() +
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold"),
        plot.title = element_markdown(),
        plot.title.position = "plot",
        axis.text.x = element_markdown(angle = 30),
        axis.text.y = element_markdown()) -> figure

#ggsave("figures/Combined1.png", width = 8, height = 4)
```

```{r eval = FALSE, include=FALSE}
(wrap_elements(full = curv1 | curv2)) / wrap_elements(full = figure) +
   plot_layout(heights = c(1.5, 1.5)) + plot_annotation(tag_levels = 'A')

#ggsave("figures/Combined2.png", height = 6)
```
